Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-08-26 manuscript/manuscript.Rmd
small updates for MEE R1

@@

\begin{enumerate}[label=\textbf{\arabic*}]
        \item Rapid environmental change has generated growing interest in forecasts of future population trajectories. Traditional population models built with detailed demographic observations from one study site can address the impacts of environmental change at particular locations, but are difficult to scale up to the landscape and regional scales relevant to management decisions. An alternative is to build models using population-level data which are much easier to collect than individual-level data over broad spatial scales. However, it is unknown whether models built using population-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.
        \item Here, we test the consequences of aggregating individual responses when forecasting the population states and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA. We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting skill and forecast horizons with and without the inclusion of climate covariates. For both models we used Bayesian ridge regression to weight the influence of climate covariates for optimal prediction. 
        \item In the absence of climate effects, we found no significant difference between the forecasting skill of models based on individual-level data and models based on population-level data. Climate effects were weak, but increased forecasting skill for two species. Increases in skill with climate covariates were similar between model types.
        \item In our case study, percent cover models generated forecasts as skillful as those from a demographic model. For the goal of forecasting, models based on aggregated individual-level data may offer a practical alternative to data-intensive demographic models. Long time series of percent cover data already exist for many plant species. Modelers should exploit these data to predict the impacts of environmental change.
    \end{enumerate}

@@

To date, most empirical population modelers have relied on individual-level data, with few attempts to capitalize on population-level measures. An important exception was an effort by @Taylor2004 to model the population growth rate of an invasive species to identify the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Such models do not require individual-level demographic data and can adequately describe population dynamics. Building on @Taylor2004, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 provide empirical evidence that density-structured models are capable of reproducing population dynamics at landscape spatial scales, even if some precision is lost when compared to fully continuous models. However, previous tests of density-structured models have yet to assess their ability to forecast out-of-sample observations, and they have not included environmental covariates, which are necessary to forecast population responses to climate change.

@@

We compared the forecasting skill of statistical and population models based on aggregated, population-level data with the skill of models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on population-level (basal cover) data. We simulated the models to answer two questions motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b].
First, can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
Second, can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data?

Materials and Methods
---------------------
### Study site and data
Our demographic data come from a northern mixed grass prairie at the Fort Keogh Livestock and Range Research Laboratory near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *Poa secunda* are cool-season perennial grasses. The growing season begins in early spring (typically in April) and lasts through mid-summer (typically in June).

@@

In equation 3, $w_{i,t}$ is the crowding that genet _i_ in year _t_ experiences from conspecific neighbors in quadrat *q*. The spatial scale over which conspecific neighbors exert influence on any genet is determined by $\delta$. The function is applied for all _k_ conspecific genets that neighbor the focal genet at time _t_, and $d_{ik,q,t}$ is the distance between genet _i_ and conspecific genet _k_ in quadrat *q*. We use regression-specific (survival and growth) $\delta$ values estimated by @Chu2015.

@@

This cross-validation procedure allowed us to compare accuracy and precision of the two modeling approaches (IPM versus QBM) with and without climate covariates. We first calculated the median predicted cover across the 100 simulations for each quadrat-year and then calculated forecast skill as the correlation ($\rho$) between forecasts and observations. We calculated forecast error as mean absolute error (MAE) between forecasts and observations. We compared $\rho$ and MAE between model types and within model types between models with and without climate covariates using one-sided *t* tests with adjusted degrees of freedom following @Wilcox2009 and standard errors calculated using the HC4 estimator of @Cribari-Neto2004. Statistical tests were conducted using algorithms from @Ye2015a.

@@

We also expected IPM forecast accuracy to decline at a lower rate than the QBM as the forecast horizon increased.
In principle, more mechanistic models should produce better predictions, especially under novel conditions [@Evans2012b; @Schindler2015b].
In our case, the IPM explicitly models the influence of weather on recruitment and survival, effects that may be poorly represented in the QBM because recruitment and survival mainly affect small plants that contribute little to year-to-year changes in percent cover.
Over time, of course, the addition and subtraction of small plants can have large effects on population growth, so explicitly modeling these effects could contribute to a longer forecast horizon. 
However, we found no evidence for a difference between the IPM and QBM forecast horizons (Fig. 4).

@@

Figures
-------
```{r figure_1, dependson="plot-options", fig.cap="Time series of average percent cover over all quadrats for our four focal species: *Bouteloua gracilis* (BOGR), *Hesperostipa comata* (HECO), *Pascopyrum smithii* (PASM), and *Poa secunda* (POSE). Light grey lines show trajectories of individual quadrats. Note the different y-axis scales across panels.", fig.height=2, fig.width=8.5, eval=T}
library(ggthemes)
obs_data <- read.csv("../analysis/data_processing/speciesData/quadAllCover.csv")
#remove zeros for averaging
obs_data <- obs_data[which(obs_data$propCover>0),]
avgobs <- ddply(obs_data, .(year, Species), summarise,
                mean_cover = mean(propCover))

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{./components/forecast_horizon.png}
  \caption{The forecast horizons for both models with climate covariates included \textcolor{blue}{and using mean parameter values. Points show the average accuracy ($\rho$) across all forecasts at a given distance between the last observation and the forecast, where forecasts are made for in-sample data.} We only examine the forecast accuracy of models with climate covariates included because in no case did including climate covariates significantly decrease accuracy (see Fig. 2). \textcolor{blue}{The dashed blue line indicates a forecast proficiency threshold of $\rho = 0.5$. Species codes are as in Fig. 1 and statistical comparisons between the IPM and QBM at each forecast distance are in Tables S23 and S24.}}
\end{figure}

>>>>

\textbf{Running head:} Demographic data and population forecasting

\textbf{Word count:} 6845 (including reference section)

\textbf{Corresponding author:}  
Andrew Tredennick (atredenn@gmail.com)  
Department of Wildland Resources and the Ecology Center  
5230 Old Main Hill  
Utah State University  
Logan, Utah 84322, USA

@@

\begin{enumerate}[label=\textbf{\arabic*}]
        \item Rapid environmental change has generated growing interest in forecasts of future population trajectories. Traditional population models built with detailed demographic observations from one study site can address the impacts of environmental change at particular locations, but are difficult to scale up to the landscape and regional scales relevant to management decisions. An alternative is to build models using population-level data which are much easier to collect than individual-level data over broad spatial scales. However, it is unknown whether models built using population-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.
        \item Here, we test the consequences of aggregating individual responses when forecasting the population states \textcolor{blue}{(percent cover)} and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA. We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting \textcolor{blue}{accuracy} and forecast horizons with and without the inclusion of climate covariates. For both models we used Bayesian ridge regression to weight the influence of climate covariates for optimal prediction. 
        \item In the absence of climate effects, we found no significant difference between the forecast \textcolor{blue}{accuracy} of models based on individual-level data and models based on population-level data. Climate effects were weak, but increased forecast \textcolor{blue}{accuracy} for two species. Increases in \textcolor{blue}{accuracy} with climate covariates were similar between model types.
        \item In our case study, percent cover models generated forecasts as \textcolor{blue}{accurate} as those from a demographic model. For the goal of forecasting, models based on aggregated individual-level data may offer a practical alternative to data-intensive demographic models. Long time series of percent cover data already exist for many plant species. Modelers should exploit these data to predict the impacts of environmental change.
    \end{enumerate}

@@

To date, most empirical population modelers have relied on individual-level data, with few attempts to capitalize on population-level measures. An important exception was an effort by @Taylor2004 to model the population growth rate of an invasive species to identify the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Such models do not require individual-level demographic data and can adequately describe population dynamics. Building on @Taylor2004, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 provide empirical evidence that density-structured models are capable of reproducing population dynamics at landscape spatial scales, even if some precision is lost when compared to fully continuous models. \textcolor{blue}{Density-structured models based on animal counts have also begun to find a home in dynamic species distribution models} [@Mieszkowska2013; @Hefley2016]. However, previous tests of density-structured population models have yet to assess their ability to forecast out-of-sample observations, and they have not included environmental covariates, which are necessary to forecast population responses to climate change.

@@

We compared the forecasting skill \textcolor{blue}{(accuracy and precision)} of statistical and population models based on aggregated, population-level data with the skill of models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on population-level (basal cover) data. We simulated the models to answer two questions motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b].
First, can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
Second, can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data?

Materials and Methods
---------------------
### Overview of analysis
\textcolor{blue}{Below we describe our focal plant populations, the population models we use to make predictions, and how we compare model forecasts, but offer a brief overview here.
We used two types of data: individual-level data and percent cover data.
Using the individual-level data we fit three vital rate regressions (survival, growth, and rectruitment) to build an Integral Projection Model (IPM) for simulating the plant populations.
Using the percent cover data we fit a simple, Gompertz-style population growth model, which we refer to as a quadrat-based model (QBM).
Both model types (IPM and QBM) we fit and simulate versions with and without climate covariates.
We used Bayesian ridge regression to weight the importance of each climate covariate.
We then performed cross-validation to assess each model's ability to predict out-of-sample data.
We compared the forecast accuracy ($\rho$, correlation between observations and predictions) and mean absolute error (MAE) between the IPM and the QBM to test our expectation that the IPM should outperform the QBM.
Lastly, we use in-sample forecasts to quantify each model's forecast horizon} [@Petchey2015].

### Study site and data
Our demographic data come from a northern mixed grass prairie at the Fort Keogh Livestock and Range Research Laboratory near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1 and Table 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *Poa secunda* are cool-season perennial grasses. The growing season begins in early spring (typically in April) and lasts through mid-summer (typically in June).

@@

In equation 3, $w_{i,q,t}$ is the crowding that genet _i_ in year _t_ experiences from _k_ conspecific neighbors ($u_{k,q,t}$) in quadrat *q*. The spatial scale over which conspecific neighbors exert influence on any genet is determined by $\delta$. The function is applied for all _k_ conspecific genets that neighbor the focal genet at time _t_, and $d_{ik,q,t}$ is the distance between genet _i_ and conspecific genet _k_ in quadrat *q*. We use regression-specific (survival and growth) $\delta$ values estimated by @Chu2015.

@@

This cross-validation procedure allowed us to compare the accuracy and precision of the two modeling approaches (IPM versus QBM) with and without climate covariates. We first calculated the median predicted cover across the 100 simulations for each quadrat-year and then calculated forecast skill as the correlation ($\rho$) between forecasts and observations. We calculated forecast error as mean absolute error (MAE) between forecasts and observations. We compared $\rho$ and MAE between model types and within model types between models with and without climate covariates using one-sided *t* tests with adjusted degrees of freedom following @Wilcox2009 and standard errors calculated using the HC4 estimator of @Cribari-Neto2004. Statistical tests \textcolor{blue}{for comparing correlations and error} were conducted using algorithms from @Ye2015a.

@@

We also expected IPM forecast accuracy to decline at a lower rate than the QBM as the \textcolor{blue}{time between the last observation and the forecast increased.}
In principle, more mechanistic models should produce better predictions, especially under novel conditions [@Evans2012b; @Schindler2015b].
In our case, the IPM explicitly models the influence of weather on recruitment and survival, effects that may be poorly represented in the QBM because recruitment and survival mainly affect small plants that contribute little to year-to-year changes in percent cover.
Over time, of course, the addition and subtraction of small plants can have large effects on population growth, so explicitly modeling these effects could contribute to a longer forecast horizon. 
However, we found no evidence \textcolor{blue}{that the forecast horizon for the IPM was greater than the QBM (Fig. 4).
On the contrary, the QBM tended to have a slightly longer forecast horizon than the IPM for most species (Fig. 4).
The QBM has fewer processes, so may better capture near term dynamics when populations do not fluctuate widely, as in our case.
When populations do not fluctuate widely, often the best forecast for population size at \emph{t+1} is the size at time \emph{t}.
Thus, a simpler model such as the QBM may be well-suited for predicting populations that fluctuate mildly around an equilibrium.}

@@

Figures
-------
```{r figure_1, dependson="plot-options", fig.cap="Time series of average percent cover over all quadrats for our four focal species: *Bouteloua gracilis* (BOGR), *Hesperostipa comata* (HECO), *Pascopyrum smithii* (PASM), and *Poa secunda* (POSE). Light grey lines show trajectories of individual quadrats. Note the different y-axis scales across panels. See Table 1 for sample size information.", fig.height=2, fig.width=8.5, eval=T}
library(ggthemes)
obs_data <- read.csv("../analysis/data_processing/speciesData/quadAllCover.csv")
#remove zeros for averaging
obs_data <- obs_data[which(obs_data$propCover>0),]
avgobs <- ddply(obs_data, .(year, Species), summarise,
                mean_cover = mean(propCover))

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{./components/forecast_horizon.png}
  \caption{The forecast horizons for both models with climate covariates included \textcolor{blue}{and using mean parameter values. Points show the average accuracy ($\rho$, \textcolor{blue}{correlation between observations and predictions}) across all forecasts at a given distance between the last observation and the forecast, where forecasts are made for in-sample data.} We only examine the forecast accuracy of models with climate covariates included because in no case did including climate covariates significantly decrease accuracy (see Fig. 2). \textcolor{blue}{The dashed blue line indicates a forecast proficiency threshold of $\rho = 0.5$. Species codes are as in Fig. 1 and statistical comparisons between the IPM and QBM at each forecast distance are in Tables S23 and S24.}}
\end{figure}

### 2016-08-26 manuscript/manuscript.tex
small updates for MEE R1

@@

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual-level demographic data and can
adequately describe population dynamics. Building on Taylor and Hastings
(2004), Freckleton et al. (2011) showed that density-structured models
compare well to continuous models in theory, and Queenborough et al.
(2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics at landscape spatial scales,
even if some precision is lost when compared to fully continuous models.
However, previous tests of density-structured models have yet to assess
their ability to forecast out-of-sample observations, and they have not
included environmental covariates, which are necessary to forecast
population responses to climate change.

@@

We compared the forecasting skill of statistical and population models
based on aggregated, population-level data with the skill of models
based on individual-level data. We used a demographic dataset that
tracks the fates of individual plants from four species over 14 years to
build two kinds of single-species population models, traditional models
using individual growth, survival, and recruitment data and alternative
models based on population-level (basal cover) data. We simulated the
models to answer two questions motivated by the fact that the effects of
intraspecific competition (density dependence) and interannual weather
variability act at the level of the individual (Clark et al. 2011).
First, can population models fit using aggregated individual-level data
(percent cover) adequately capture density dependence to produce
forecasts as skillful as those from models fit to demographic data?
Second, can population models fit using aggregated data adequately
capture the influence of climate on population growth and, in turn,
produce forecasts as skillful as those from models fit to demographic
data?

@@

@@

Our demographic data come from a northern mixed grass prairie at the
Fort Keogh Livestock and Range Research Laboratory near Miles City,
Montana, USA (\(46^{\circ}\) 19' N, \(105^{\circ}\) 48' W). The dataset
is available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011), and interested readers should refer to the
metadata for a complete description. The site is 800 m above sea level
and mean annual precipitation (1878-2009) is 334 mm, with most annual
precipitation falling from April through September. The community is
grass-dominated, and we focused on the four most abundant grass species:
\emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO),
\emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig.
1). \emph{B. gracilis} is a warm-season perennial grass, whereas
\emph{H. comata}, \emph{P. smithii}, and \emph{Poa secunda} are
cool-season perennial grasses. The growing season begins in early spring
(typically in April) and lasts through mid-summer (typically in June).

@@

In equation 3, \(w_{i,t}\) is the crowding that genet \emph{i} in year
\emph{t} experiences from conspecific neighbors in quadrat \emph{q}. The
spatial scale over which conspecific neighbors exert influence on any
genet is determined by \(\delta\). The function is applied for all
\emph{k} conspecific genets that neighbor the focal genet at time
\emph{t}, and \(d_{ik,q,t}\) is the distance between genet \emph{i} and
conspecific genet \emph{k} in quadrat \emph{q}. We use
regression-specific (survival and growth) \(\delta\) values estimated by
Chu and Adler (2015).

@@

This cross-validation procedure allowed us to compare accuracy and
precision of the two modeling approaches (IPM versus QBM) with and
without climate covariates. We first calculated the median predicted
cover across the 100 simulations for each quadrat-year and then
calculated forecast skill as the correlation (\(\rho\)) between
forecasts and observations. We calculated forecast error as mean
absolute error (MAE) between forecasts and observations. We compared
\(\rho\) and MAE between model types and within model types between
models with and without climate covariates using one-sided \emph{t}
tests with adjusted degrees of freedom following Wilcox (2009) and
standard errors calculated using the HC4 estimator of Cribari-Neto
(2004). Statistical tests were conducted using algorithms from Ye et al.
(2015).

@@

We also expected IPM forecast accuracy to decline at a lower rate than
the QBM as the forecast horizon increased. In principle, more
mechanistic models should produce better predictions, especially under
novel conditions (Evans 2012, Schindler and Hilborn 2015). In our case,
the IPM explicitly models the influence of weather on recruitment and
survival, effects that may be poorly represented in the QBM because
recruitment and survival mainly affect small plants that contribute
little to year-to-year changes in percent cover. Over time, of course,
the addition and subtraction of small plants can have large effects on
population growth, so explicitly modeling these effects could contribute
to a longer forecast horizon. However, we found no evidence for a
difference between the IPM and QBM forecast horizons (Fig. 4).

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1-1.pdf}
\caption{Time series of average percent cover over all quadrats for our
four focal species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa
comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda}
(POSE). Light grey lines show trajectories of individual quadrats. Note
the different y-axis scales across panels.}
\end{figure}

@@

@@

>>>>

\textbf{Running head:} Demographic data and population forecasting

\textbf{Word count:} 6845 (including reference section)

\textbf{Corresponding author:}\\Andrew Tredennick
(\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}})\\Department
of Wildland Resources and the Ecology Center\\5230 Old Main Hill\\Utah
State University\\Logan, Utah 84322, USA

@@

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual-level demographic data and can
adequately describe population dynamics. Building on Taylor and Hastings
(2004), Freckleton et al. (2011) showed that density-structured models
compare well to continuous models in theory, and Queenborough et al.
(2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics at landscape spatial scales,
even if some precision is lost when compared to fully continuous models.
\textcolor{blue}{Density-structured models based on animal counts have also begun to find a home in dynamic species distribution models}
(Mieszkowska et al. 2013, Hefley and Hooten 2016). However, previous
tests of density-structured population models have yet to assess their
ability to forecast out-of-sample observations, and they have not
included environmental covariates, which are necessary to forecast
population responses to climate change.

@@

We compared the forecasting skill
\textcolor{blue}{(accuracy and precision)} of statistical and population
models based on aggregated, population-level data with the skill of
models based on individual-level data. We used a demographic dataset
that tracks the fates of individual plants from four species over 14
years to build two kinds of single-species population models,
traditional models using individual growth, survival, and recruitment
data and alternative models based on population-level (basal cover)
data. We simulated the models to answer two questions motivated by the
fact that the effects of intraspecific competition (density dependence)
and interannual weather variability act at the level of the individual
(Clark et al. 2011). First, can population models fit using aggregated
individual-level data (percent cover) adequately capture density
dependence to produce forecasts as skillful as those from models fit to
demographic data? Second, can population models fit using aggregated
data adequately capture the influence of climate on population growth
and, in turn, produce forecasts as skillful as those from models fit to
demographic data?

@@

\subsubsection{Overview of analysis}\label{overview-of-analysis}

\textcolor{blue}{Below we describe our focal plant populations, the population models we use to make predictions, and how we compare model forecasts, but offer a brief overview here.
We used two types of data: individual-level data and percent cover data.
Using the individual-level data we fit three vital rate regressions (survival, growth, and rectruitment) to build an Integral Projection Model (IPM) for simulating the plant populations.
Using the percent cover data we fit a simple, Gompertz-style population growth model, which we refer to as a quadrat-based model (QBM).
Both model types (IPM and QBM) we fit and simulate versions with and without climate covariates.
We used Bayesian ridge regression to weight the importance of each climate covariate.
We then performed cross-validation to assess each model's ability to predict out-of-sample data.
We compared the forecast accuracy ($\rho$, correlation between observations and predictions) and mean absolute error (MAE) between the IPM and the QBM to test our expectation that the IPM should outperform the QBM.
Lastly, we use in-sample forecasts to quantify each model's forecast horizon}
(Petchey et al. 2015).

@@

Our demographic data come from a northern mixed grass prairie at the
Fort Keogh Livestock and Range Research Laboratory near Miles City,
Montana, USA (\(46^{\circ}\) 19' N, \(105^{\circ}\) 48' W). The dataset
is available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011), and interested readers should refer to the
metadata for a complete description. The site is 800 m above sea level
and mean annual precipitation (1878-2009) is 334 mm, with most annual
precipitation falling from April through September. The community is
grass-dominated, and we focused on the four most abundant grass species:
\emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO),
\emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1
and Table 1). \emph{B. gracilis} is a warm-season perennial grass,
whereas \emph{H. comata}, \emph{P. smithii}, and \emph{Poa secunda} are
cool-season perennial grasses. The growing season begins in early spring
(typically in April) and lasts through mid-summer (typically in June).

@@

In equation 3, \(w_{i,q,t}\) is the crowding that genet \emph{i} in year
\emph{t} experiences from \emph{k} conspecific neighbors (\(u_{k,q,t}\))
in quadrat \emph{q}. The spatial scale over which conspecific neighbors
exert influence on any genet is determined by \(\delta\). The function
is applied for all \emph{k} conspecific genets that neighbor the focal
genet at time \emph{t}, and \(d_{ik,q,t}\) is the distance between genet
\emph{i} and conspecific genet \emph{k} in quadrat \emph{q}. We use
regression-specific (survival and growth) \(\delta\) values estimated by
Chu and Adler (2015).

@@

This cross-validation procedure allowed us to compare the accuracy and
precision of the two modeling approaches (IPM versus QBM) with and
without climate covariates. We first calculated the median predicted
cover across the 100 simulations for each quadrat-year and then
calculated forecast skill as the correlation (\(\rho\)) between
forecasts and observations. We calculated forecast error as mean
absolute error (MAE) between forecasts and observations. We compared
\(\rho\) and MAE between model types and within model types between
models with and without climate covariates using one-sided \emph{t}
tests with adjusted degrees of freedom following Wilcox (2009) and
standard errors calculated using the HC4 estimator of Cribari-Neto
(2004). Statistical tests
\textcolor{blue}{for comparing correlations and error} were conducted
using algorithms from Ye et al. (2015).

@@

We also expected IPM forecast accuracy to decline at a lower rate than
the QBM as the
\textcolor{blue}{time between the last observation and the forecast increased.}
In principle, more mechanistic models should produce better predictions,
especially under novel conditions (Evans 2012, Schindler and Hilborn
2015). In our case, the IPM explicitly models the influence of weather
on recruitment and survival, effects that may be poorly represented in
the QBM because recruitment and survival mainly affect small plants that
contribute little to year-to-year changes in percent cover. Over time,
of course, the addition and subtraction of small plants can have large
effects on population growth, so explicitly modeling these effects could
contribute to a longer forecast horizon. However, we found no evidence
\textcolor{blue}{that the forecast horizon for the IPM was greater than the QBM (Fig. 4).
On the contrary, the QBM tended to have a slightly longer forecast horizon than the IPM for most species (Fig. 4).
The QBM has fewer processes, so may better capture near term dynamics when populations do not fluctuate widely, as in our case.
When populations do not fluctuate widely, often the best forecast for population size at \emph{t+1} is the size at time \emph{t}.
Thus, a simpler model such as the QBM may be well-suited for predicting populations that fluctuate mildly around an equilibrium.}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1-1.pdf}
\caption{Time series of average percent cover over all quadrats for our
four focal species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa
comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda}
(POSE). Light grey lines show trajectories of individual quadrats. Note
the different y-axis scales across panels. See Table 1 for sample size
information.}
\end{figure}

@@

Hefley, T. J., and M. B. Hooten. 2016. Hierarchical Species Distribution
Models. Current Landscape Ecology Reports 1:87--97.

@@

Mieszkowska, N., G. Milligan, M. T. Burrows, R. Freckleton, and M.
Spencer. 2013. Dynamic species distribution models from categorical
survey data. Journal of Animal Ecology 82:1215--1226.
